---
title: "Lab 02 - Plastic waste"
author: "Cat Seitz"
date: "01/16/23"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
library(viridis)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1


```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```
```{r filter to find highest}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

I wasn't expecting this result, but upon further investigation, I found that Trinidad and Tobago has inefficient waste management systems and the households don't segregate waste. 


```{r plot dist of plastic waste per capita}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_histogram(alpha = 0.7)
```

### Exercise 2.1

```{r plastic-waste-density, warning=FALSE}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_histogram(alpha = 0.3)
```

### Exercise 2.2

The color and fill are defined by the continent so the dataframe gets grouped that way instead of all of the continents grouped together, but alpha is a function of the entire graph. 
```{r box plots by continent}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```


### Exercise 3.1

Violin plots show the density of the distribution as well as all of the elements of the box plot. 

```{r violin plots}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 4.1

```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap,
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```
There seems to be a positive correlation between between plastic waste per capita and mismanaged waste per capita. 


### Exercise 4.2

```{r plastic-waste-mismanaged by continent}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap,
                     y = mismanaged_plastic_waste_per_cap, 
                     color=continent))+
  geom_point()
```
It's somewhat hard to tell because most of the data points are confined to the same area, but it looks like some continents have steeper relationships between plastic waste per capita and mismanaged waste per capita. If I has to speculate, Europe has a less steep relationship and Africa has a more steep relationship. 


### Exercise 4.3

```{r plastic-waste by total pop}
ggplot(data = plastic_waste, 
       mapping = aes(x = total_pop,
                     y = plastic_waste_per_cap))+
  geom_point()
```
```{r plastic-waste by coastalpop}
ggplot(data = plastic_waste, 
       mapping = aes(x = coastal_pop,
                     y = plastic_waste_per_cap))+
  geom_point()
```

Plastic waste per capita and coastal population seem to be more linearly associated than plastic waste per capita and total population, but, again, it's difficult to tell just from these scatterplots. 



### Exercise recreate plot

```{r recreate-viz}
plastic_waste <- transform(
  plastic_waste, coastal_pop_proportion= plastic_waste$coastal_pop/plastic_waste$total_pop
)
```

```{r graph it}

ggplot(plastic_waste[which(plastic_waste$plastic_waste_per_cap<3.5),],
       mapping = aes(x = coastal_pop_proportion,
                     y = plastic_waste_per_cap))+
  geom_point(aes(color=continent))+
  geom_smooth(color="#000000")+
  scale_color_viridis(discrete=TRUE)+
  
  labs(title = "Plastic waste vs. coastal population proportion", subtitle = "by continent")+
  labs(x = "Coastal population proportion (Coastal/total population)", y = "Plastic waste per capita")

```


