---
title: "Lab 02 - Plastic waste"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```
```{r filter to find highest}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
I wasn't expecting this result, but upon further investigation, I found that Trinidad and Tobago has inefficient waste management systems and the households don't segregate waste. 


```{r plot dist of plastic waste per capita}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_histogram(alpha = 0.7)
```

### Exercise 2.1

```{r plastic-waste-density}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_histogram(alpha = 0.3)
```
### Exercise 2.2

The color and fill are defined by the continent so the dataframe gets grouped that way instead of all of the continents grouped together, but alpha is a function of the entire graph. 
```{r box plots by continent}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```


### Exercise 3.1

Violin plots show the density of the distribution as well as all of the elements of the box plot. 

```{r violin plots}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 4

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.



### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



